---
title: "STAT 1060 Data Science Foundations"
subtitle: "Homework 9"
#author: "Instructor Version"
#date: "8/12/2022"
output: 
    html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.

#### Questions 1-3: Exercises 1-3 in Lab 9

Import the Titanic data set and check some basic information about it:

```{r}
titanicData <- read.csv("/Users/SanjanaPejathaya/Downloads/titanic.csv", stringsAsFactors = TRUE)
dim(titanicData)
head(titanicData,3)
colnames(titanicData)
```


##### <span style="color:purple"> Exercise 1. (10 points) Side-by-Side Boxplot </span>
Suppose we are interested in comparing the mean ages among different passenger classes.

Create a side-by-side boxplot to compare the age distributions among passenger classes. 

Note that boxplots will not show the means. Instead, only the medians will be displayed. Remember, in ANOVA, our goal is the compare means among groups. So, make sure that you mark the sample means in the boxplot.

```{r}
attach(titanicData)
library(ggplot2)
ggplot(titanicData, aes(x = passenger_class, y = age, fill = passenger_class)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=18, size=3, color="red", position=position_dodge(0.75)) +
  labs(title = "Age Distribution Among Passenger Classes",
       x = "Passenger Class",
       y = "Age") +
  theme_minimal()

```

##### <span style="color:purple"> Exercise 2.  (10 points) ANOVA </span>
Suppose you are interested in comparing the mean age among different groups of `embarked`. Note that there are only three locations to embark:

```{r}
unique(embarked)

model <- lm(age ~ embarked, data = titanicData)


anova_result <- anova(model)
print(anova_result)


summary(model)


```
Perform an ANOVA test to test the following hypotheses:

* $H_0$: Mean ages are all the same among all `embarked` groups.
* $H_a$: Mean ages are not all the same among all `embarked` groups.

What is your conclusion?

The p-value is very small and therefore we can reject the null hypotheses of no differences in mean age of all the embarked groups. 

##### <span style="color:purple"> Exercise 3.  (10 points) Permutation test for ANOVA </span>
Perform a permutation test for the hypotheses in Exercise 2. What is your conclusion?

```{r}
library(lmPerm)
anova_perm <- aovp(age ~ embarked, data = titanicData)
summary(anova_perm)
```

##### <span style="color:purple"> Question 4  (20 points) Understanding the ANOVA Table </span>
It is estimated that more than 90% of North American students use social media. In one study, researchers investigated how mental health workers perceive the association between social media and mental disorders. A sample of psychiatrists from Canada completed a questionnaire from which a perception score was obtained (a higher score indicating a stronger perceived association).

The following (incomplete) ANOVA table summarizes a comparison of these scores across three age groups (generations).

```{r, echo=FALSE}
library(kableExtra)
df <- data.frame(Source = c("Age", "Error", "Total"), 
                 DF = c(NA, 45, 47),
                 SS = c(137.78, NA, 7037.32),
                 MS = c(NA, 153.32, NA),
                 F_statistic = c(NA, NA, NA))
opts <- options(knitr.kable.NA = "")
df %>% 
  kbl() %>% 
  kable_styling(full_width = F) %>% 
  column_spec(c(2,3,4,5), width = "5em")


```

1. (10 points) Complete the above ANOVA table. 

```{r, echo = T}
library(kableExtra)
library(dplyr)

# Given data
df <- data.frame(Source = c("Age", "Error", "Total"),
                 DF = c(NA, 45, 47),
                 SS = c(137.78, NA, 7037.32),
                 MS = c(NA, 153.32, NA),
                 F_statistic = c(NA, NA, NA))

# Calculate missing values
df$DF[1] <- 1  # Degrees of Freedom for Age
df$MS[1] <- df$SS[1] / df$DF[1]  # Mean Squares for Age
df$MS[3] <- df$SS[3] / df$DF[2]  # Mean Squares for Total (assuming SS_Total = SS_Age + SS_Error)
df$SS[2] <- df$MS[2] * df$DF[2]  # Sum of Squares for Error
df$F_statistic[1] <- df$MS[1] / df$MS[2]  # F-statistic for Age
df$F_statistic[2] <- df$MS[2] / df$MS[3]  # F-statistic for Error
df$F_statistic[3] <- df$MS[3]/ df$MS[1]

# Copied from the code chunk provided 
df %>%
  kbl() %>%
  kable_styling(full_width = F) %>%
  column_spec(c(2, 3, 4, 5), width = "5em")



```


2. (10 points) Answer the following questions:
    (a) (2 points) How many psychiatrists completed the questionnaire?
    (b) (3 points) What is the estimated common standard deviation?
    (c) (3 points) What is the p-value? 
    (d) (2 points) State your conclusion using the p-value from part (c) and a 5% significance
    level.
    
```{r, echo = TRUE}

#(a) Code

DF_Error <- 45
DF_Age <- 1  

total_number_of_obs <- DF_Error + DF_Age + 1

print(total_number_of_obs)

#(b) Code

SS_Within <- 1
df_Within <- 45

pooled_sd <- sqrt(SS_Within / df_Within)

print(pooled_sd)

#(c) Code

F_statistic <- 4.3  
df_Between <- 2    
df_Within <- 45     


p_value <- pf(F_statistic, df_Between, df_Within, lower.tail = FALSE)

print(p_value)

# (d) since the p_value < significance level. Therefore we can reject the null hypotheses and state that there is a difference between the means of this data set. 


```