#### Questions 1-6: Collect your solutions to Exercises 1-6 in Lab 3. 

We will conduct an exploratory data analysis using the WHO life expectancy data set, which contains data of representative health-related factors from 2000 to 2015 for 193 countries. 

Read and attach the data set:
```{r}
LifeExp = read.csv("LifeExpectancy.csv",header=T)
attach(LifeExp)
```

##### <span style="color:purple"> Exercise 1.(5 points) </span>
Create a histogram for `Schooling` and choose appropriete set of summary statistics for it (i.e. measures of center and variability).

```{r, echo = TRUE}
hist(Life.expectancy)
median(Life.expectancy, na.rm = TRUE)
mean(Life.expectancy, na.rm = TRUE)
```

##### <span style="color:purple"> Exercise 2. (5 points)  </span>
Create a horizontal boxplot for `Schooling` and calculate the 5-number summary for it. Compare the plot with the numeric summary.

``` {r, echo= TRUE}

boxplot(Schooling, horizontal = TRUE, main = "Schooling")
quantile(Schooling, na.rm= TRUE)

```
##### <span style="color:purple"> Exercise 3. (5 points) </span>
Create a pie chart for `Status`.

```{r, echo = TRUE}
tb_status <- table(Status)
pie(tb_status)

```
##### <span style="color:purple"> Exercise 4. (5 points) </span>
Create a scatterplot for `Life.expectancy` and `Alcohol`. Do you think we can use correlation to quantify the relationship between them? If so, calculate the correlation between them.

```{r, echo = TRUE}
plot(x=Life.expectancy, y=Alcohol)
correlation <- cor(x=Life.expectancy, y=Alcohol, use='complete.obs')
print(correlation)

```

##### <span style="color:purple"> Exercise 5. (5 points) </span>
Create a scatterplot for `Life.expectancy` and `GDP`. Do you think we can use correlation to quantify the relationship between them? If so, calculate the correlation between them.


``` {r, echo = TRUE}

plot(x=Life.expectancy, y=GDP, )
correlation <- cor(x=Life.expectancy, y=GDP, use='complete.obs')
print(correlation)

```

##### <span style="color:purple"> Exercise 6. (5 points) </span>
Modify the last boxplot in lab 3 according the requirements as follows:

1. Add a title "Life Expectancy for Developed and Developing Countries".
2. Change the x-axis label to "Life Expectancy (year)".
3. Change the color of the boxplot to light blue.

``` {r, echo = TRUE}

boxplot(Life.expectancy~Status, horizontal=T, 
        main="Life Expectancy for Developed and Developing Countries",
        xlab="Life Expectancy (year)",
        col="lightblue")

```
##### <span style="color:purple"> Question 7 (4 points)</span>
All human blood can be "ABO-typed" as one of O, A, B, or AB, but the distribution of the types varies a bit among groups of people. Here is the disbribution of blood types for a randomly chosen person in the United States:


|Blood type | A | B | AB | O |
|:--|:--:|:--:|:--:|:--:|
|U.S. probability| 0.42 | ? | 0.03 | 0.44|


(a) (2 points) What is the probability of type B in the United States?

``` {r, echo= TRUE}
1- 0.42-0.03-0.44

```



(b) (2 points) Maria has type B blood. She can safely receive blood transfusions from people with blood types O and B. What is the probability that a randomly chosen person from the United States can donate blood to Maria?

``` {r, echo= TRUE}
0.11 + 0.44

```


##### <span style="color:purple"> Question 8 (6 points)</span>
The distribution of blood types in Ireland differs from the U.S. distribution given in the previous exercise:

|Blood type | A | B | AB | O |
|:--|:--:|:--:|:--:|:--:|
|Ireland probability| 0.35 | 0.10 | 0.03 | 0.52|

(a) (3 points) Choose a person from the U.S. and a person from Ireland at random, independently of each other. What is the probability that both have type O blood?

```{r, echo= TRUE}
# P(A and B)

0.44 * 0.52

```

(b) (3 points) What is the probability that both have the same blood?

```{r, echo = TRUE}
# P(A) + P(B) + P(AB) + P(O)

(0.42 * 0.35) + (0.11 * 0.10) + (0.03 * 0.03) + (0.44 * 0.52)

```


##### <span style="color:purple"> Question 9 (2 points)</span>
The probability that an event A will occur is 0.138, and the probability that B will occur is 0.261. The probability that both A and B occur is 0.082. What is the probability that either A or B occurs?

```{r, echo = TRUE}

# P(A U B) = P(A) + P(B) - P(AB)

0.138 + 0.082 - (0.138*0.082)

```


##### <span style="color:purple"> Question 10 (8 points)</span>
In the population of young children eligible to participate in a study of whether or not their calcium intake is adequate, 52\% are 5 to 10 years of age and 48\% are 11 to 13 years of age. For those who are 5 to 10 years of age, 18\% have inadequate calcium intake. For those who are 11 to 13 years of age, 57\% have inadequate calcium intake.

(a) (4 points) Calculate the probability that a randomly selected child from this population has an inadequate intake of calcium. 

```{r, echo=TRUE}

#P(I) = P(I|A) * P(A) + P(I|B) * P(B)

P_A <- 0.52
P_B <- 0.48
P_I_given_A <- 0.18
P_I_given_B <- 0.57


P_I <- P_I_given_A * P_A + P_I_given_B * P_B


print(P_I)


```



(b) (4 points) Use Bayes' rule to find the probability that a child from this population who has inadequate intake is 11 to 13 years old.

```{r, echo = TRUE}
#P(B|I) = P(I|B) * P(B)/ P

P_B_given_I <- (P_I_given_B * P_B) / P_I

print(P_B_given_I)

```

Hint: Use R as an efficient calculator for Questions 7 to 10.

```{r}
detach(LifeExp)
```

