-c--
title: "STAT 1060 Data Science Foundations"
subtitle: "Lab 3 Exploratory Data Analysis"
author: "Instructor Version"
#date: "8/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, out.width="70%")
```

The goal of this lab is to practice skills in running descriptive statistics and creating graphs using R. Make sure you've read and fully understood lecture slides of Lectures 4 to 9 as this guide tracks closely with the material presented there.

## <span style="color:navy"> Exploratory Data Analysis </span>

We will conduct an exploratory data analysis using the WHO life expectancy data set, which contains data of representative health-related factors from 2000 to 2015 for 193 countries. 

```{r}
LifeExp = read.csv("LifeExpectancy.csv",header=T)
names(LifeExp)
```

#### <span style="color:salmon">Description of some variables</span>

- Life.Expectancy: Life Expectancy in age
- Status: Developed or developing
- Alcohol: Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)               
- percentage.expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita (%)        
- Hepatitis.B: Hepatitis B (HepB) immunization coverage among 1-year-olds (%)   
- Measles: number of reported cases per 1000 population               
- BMI: Average Body Mass Index of entire population                         
- GDP: Gross Domestic Product per capita (in USD)                         
- Population: Population of the country
- Schooling: Number of years of Schooling (in years)

#### <span style="color:salmon">Original goals of the study</span>

The data-set aims to answer the following key questions:

1. Does various predicting factors which has been chosen initially really affect life expectancy? What are the predicting variables actually affecting life expectancy?
2. Should a country having a lower life expectancy value(<65) increase its **healthcare expenditure** in order to improve its average lifespan?
3. How does Infant and Adult **mortality rates** affect life expectancy?
4. Does life expectancy has positive or negative correlation with **eating habits, lifestyle, exercise, smoking, drinking alcohol** etc.?
5. What is the impact of **schooling** on the lifespan of humans?
6. Does Life Expectancy have positive or negative relationship with drinking **alcohol**?
7. Do **densely populated** countries tend to have lower life expectancy?
8. What is the impact of **immunization coverage** on life expectancy?


#### <span style="color:salmon">Goals of our EDA</span>
As you can tell that this is a very comprehensive study. In order to be more focused on the R functions that we have learned for descriptive statistics and graphical display, we will only consider the following two questions:

1. What is the impact of `Schooling` on the lifespan of humans?
2. What is the impact of `Status` on the lifespan of humans?

## <span style="color:navy"> Exploring Single Numeric Variables </span>
For a single numeric variable, we are interested in estimating their location (center) and variability. To choose the more appropriate set of descriptive statistics, we might need to visualize the data first. 

For example, let us first create a histogram of the response variable -- `Life.Expectancy`:
```{r, out.width="50%"}
attach(LifeExp)
hist(Life.expectancy)
```

The histogram is left-skewed. So, the median is a better measure of center than the mean.

```{r}
median(Life.expectancy)
```
We got `NA` because there are missing values. To get a valid value, we can choose to ignore the missing value:

```{r}
median(Life.expectancy, na.rm=TRUE)
```
So, the median life expectancy is about 72 years. How about the mean?

```{r}
mean(Life.expectancy, na.rm=TRUE)
```
The mean is only about 69 years, because there are some extreme values around the left tail.

##### <span style="color:purple"> Exercise 1.Create a histogram for `Schooling` and choose appropriete set of summary statistics for it (i.e. measures of center and variability).</span>
```{r, out.width="50%"}
hist(Schooling)

```

```{r}
median(Schooling,na.rm=T)
mad(Schooling, na.rm=T)
```

##### <span style="color:purple"> Exercise 2. Create a horizontal boxplot for `Schooling` and calculate the 5-number summary for it. Compare the plot with the numeric summary.</span>

```{r}
boxplot(Schooling, horizontal=T)
```

```{r}
quantile(Schooling, na.rm=T)
```
## <span style="color:navy"> Exploring Single Categorical Variables </span>
Next, let us study the distribution of `Status`. This is a categorical variable with only two values: `developed` and `developing`.

Let us first create a distribution table for `Status`:
```{r}
tb_status = table(Status)
tb_status
```

To show the proportions instead of counts, we may divide the table by the total number of observations:
```{r}
tb_status/length(Status)
```

There are about 83.6\% of developing countries and 17.4\% of developed countries.

Now, let us create a bar graph to show the distribution:
```{r, out.width="50%"}
barplot(tb_status)
```

##### <span style="color:purple"> Exercise 3. Create a pie chart for `Status`.</span>

```{r, out.width="50%"}
pie(tb_status)
```

## <span style="color:navy"> Exploring Two Numeric Variables </span>
Since we are interested in the relationship between `Schooling` and `Life.expectancy`, let us first create a scatterplot for them.
```{r, out.width="50%"}
plot(x=Schooling, y=Life.expectancy)
```


It seems that there is a moderately strong linear positive relationship between these two variables. 

To quantify the linear relationship, we will calculate the correlation coefficient, $r$:
```{r}
cor(Schooling, Life.expectancy, use="complete.obs")
```
Notice that we chose to use complete observations. The correlation is 0.75, which is consistent with what we have observed from the scatterplot.

##### <span style="color:purple"> Exercise 4. Create a scatterplot for `Life.expectancy` and `Alcohol`. Do you think we can use correlation to quantify the relationship between them? If so, calculate the correlation between them.</span>


```{r, out.width="50%"}
plot(Alcohol, Life.expectancy)
```

```{r}
cor(Alcohol, Life.expectancy, use="complete.obs")
```

##### <span style="color:purple"> Exercise 5. Create a scatterplot for `Life.expectancy` and `GDP`. Do you think we can use correlation to quantify the relationship between them? If so, calculate the correlation between them.</span>

```{r, out.width="50%"}
plot(GDP, Life.expectancy)
```

The relationship is not linear, so we should not use correlation to quantify the relationship between these two variables.


## <span style="color:navy"> Exploring One Numeric Variable and One Categorical Variable </span>
Finally, let us study the relationship between `Status` and `Life.expectancy`. Of course, we cannot make conclusion based on the result of EDA. Later on, we will learn how to use hypothesis testing to quantify the evidence from data.

First, side-by-side boxplot is a good tool to compare grouped numeric data:
```{r}
boxplot(Life.expectancy~Status, horizontal=T)
```

Developed country has higher median, smaller range and interquartile range. In general, we may guess that life expectancy in developed country is higher.

##### <span style="color:purple"> Exercise 6. Modify the above boxplot according the requirements as follows:</span>

1. Add a title "Life Expectancy for Developed and Developing Countries".
2. Change the x-axis label to "Life Expectancy (year)".
3. Change the color of the boxplot to light blue.

```{r}
boxplot(Life.expectancy~Status, horizontal=T, 
        main="Life Expectancy for Developed and Developing Countries",
        xlab="Life Expectancy (year)",
        col="lightblue")
```

```{r}
detach(LifeExp)
```

