---
title: "STAT 1060 Data Science Foundations"
subtitle: "Lab 4 Normal Functions in R"
#date: "8/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, out.width="70%")
```

The goal of this lab is to explore normal functions built-in R. In Lectures 13 and 14, we have studied the normal distribution family. 

In today's lab, we will learn to use R functions to 

1. Show the density curve of a normal distribution;
2. Calculate probabilities related to normal distributions;
3. Calculate percentiles (quantiles) of a normal distribution;
4. Generate random numbers from a normal distribution;

### <span style="color:navy"> Normal Distributions </span>
Recall that a normal distribution has two parameters: $\mu$ and $\sigma$, and if a random variable follows a normal distribution, it is usually denoted by
$$X\sim N(\mu,\sigma)$$
where $\mu$ and $\sigma$ are the mean and standard deviation of the distribution of $X$, $x\in (-\infty,+\infty)$. 

#### <span style="color:salmon">Density Function</span>
The density function of $N(\mu,\sigma)$ distribution is:
$$f(x)=\frac{1}{\sqrt{2\pi}\sigma}\exp\left\{{-\frac{(x-\mu)^2}{2\sigma^2}}\right\}$$

A special case of normal distribution: Standard Normal, $Z\sim N(0,1)$, with probability density function (pdf):
$$f(z)=\frac{1}{\sqrt{2\pi}}e^{-\frac{z^2}{2}}$$

#### <span style="color:salmon">Cumulative Probability Function (cdf)</span>
The CDF of a normal distribution is defined as
$F(x)=P(X\le x) =\int_{-\infty}^{x}f(t)dt$

### <span style="color:navy"> Built-in Functions in R </span>
The standard distributions that are commonly used in model-building and statistical tests have been built in `R`, and it can therefore completely replace the traditional statistical tables.

Four fundamental items can be calculated for a statistical distribution:

- Probability density function (pdf) or probability mass function (pmf)
- Cumulative probability distribution function (cdf)
- Quantiles (percentiles)
- Pseudo-random numbers

For all distributions implemented in `R`, there is a function for each of the four items listed above.

For example, for the normal distribution, these are named `dnorm`, `pnorm`, `qnorm`,\  and `rnorm`.

| Type | R Function | Example|
|:--|:--|:--|
|density function | `dnorm(x, mean=0, sd=1)` | `dnorm(1.5)`|
|cumulative distribution function | `pnorm(q, mean=0, sd=1, lower.tail = TRUE)`| `pnorm(1.5, 100, 15)`|
|quantile function| `qnorm(p, mean=0, sd=1, lower.tail = TRUE)` | `qnorm(0.95)`|
|pseudo-random number| `rnorm(n, mean=0, sd=1)`| `rnorm(100,2,10)`|

#### <span style="color:salmon">The Normal Distribution (1): Density Function </span>

The density for a continuous distribution is a measure of the relative probability of "getting a value close to x". The probability of getting a value in a particular interval is the area under the corresponding part of the curve. 
The density function is likely the least used in practice, but if you want to draw the well-known bell curve of the normal distribution, then it can be done like this:

```{r,fig.height=4}
par(mfrow=c(1,2))
x <- seq(-4,4,0.1)
plot(x, dnorm(x), type="l", ylab="f(x)",lwd=2)
y <- seq(0,16,0.1)
plot(y, dnorm(y, mean=8, sd=2),, type="l", ylab="f(y)",lwd=2,col=2)
```

Note that a layout with two plots side by side is specified using the `mfrow` graphical parameter. It should be read as "multi-frame, rowwise, 1-by-2 layout".



#### <span style="color:salmon">The Normal Distribution (2): Cumulative Distribution Function </span>

The cumulative distribution function is defined as $F(x)=P(X\le x).$ For the normal distribution, 
$$F(x)=\int_{-\infty}^x \frac{1}{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}} dx $$

In `R`, we can use the `pnorm` function to calculate the cumulative probability. For example, find $P(Z<-2):$

```{r}
pnorm(-2)
```

##### <span style="color:purple"> Exercise 1. Calculate $P(Z>4)$ where $Z\sim N(0,1)$.</span>


Similar to `dnorm` function, it is possible to use `pnorm` function to find cumulative probability of **any** normal distribution. 

For example, suppose $X\sim N(100,15),$ find the probability that $P(X>135).$
```{r}
pnorm(135, mean = 100, sd = 15, lower.tail=FALSE)
```

##### <span style="color:purple"> Exercise 2. Calculate $P(85 <X < 115)$ where $X\sim N(100,15).$  </span>


How does the cumulative distribution function look like?
```{r,fig.height=4}
par(mfrow=c(1,2))
x <- seq(-4,4,0.1)
plot(x, pnorm(x), type="l", ylab="F(x)",lwd=2)
y <- seq(0,16,0.1)
plot(y, pnorm(y, mean=8, sd=2),, type="l", ylab="F(y)",lwd=2,col=2)
```

##### <span style="color:purple"> Exercise 3. </span>

Typical nightly hours slept by college students is a normal random variable with $\mu = 7$ and $\sigma = 1.5$. 

(a) What is the probability that a student slept less than 6 hours?

(b) What is the probability that a student slept more than 12 hours?

#### <span style="color:salmon">The Normal Distribution (3): Quantiles/Percentiles </span>

The quantile (percentile) function is the inverse of the cumulative distribution function. The **$p$-quantile** (or $p$-percentile) is the value with the property that there is probability $p$ of getting a value less than or equal to it, i.e. if $x$ is the $p$-quantile, then it must be true that
$$P(X<x)=p$$
The standard normal table contains typical $z$ values and corresponding cumulative probabilities, $p$. Actually, each of these $z$ values is the corresponding $p$-percentile. Given a cumulative probability, $p$, we can find the corresponding $z$ value in the table. In `R`, we may use `qnorm` function.

For example, if $F(z)=0.66,$ then what is the value of $z?$ This question is actually asking about the $66^{th}$-percentile of standard normal distribution.
```{r}
qnorm(0.66)
```
This implies that $P(Z<0.424631)=0.66$.

##### <span style="color:purple"> Exercise 4. Standard Normal Quantile.</span>

(a) Suppose $P(Z>z)=0.95$, what is the value of $z$?
(b) What is the $99^{th}$-percentile (or 0.99-quantile) of standard normal distribution?

#### Example: IQ
IQ for randomly chosen adult is a normal random variable $X$ with $\mu=100$ and $\sigma=15.$ In order to be among the top 5% smartest people, what IQ score one needs to have?

We are looking for an $x$ value such that $P(X>x) = 0.05,$ i.e. $P(X<x)=0.95.$ Therefore, $x$ is the $95^{th}$ percentile of $X.$

```{r}
qnorm(0.95, mean = 100, sd = 15)
```

##### <span style="color:purple"> Exercise 5. Normal Quantiles.</span>
In order to be among the top 1% smartest people, what IQ score one needs to have? 


#### <span style="color:salmon">The Normal Distribution (4): Random Numbers </span>

Although it sounds like a contradiction in terms to generate random numbers on a computer since its results are supposed to be predictable and reproducible, it is possible to generate sequence of "pseudo-random" numbers, which for practical purposes behave **as if they were drawn randomly**.

In professional statistics, these "pseudo-random" numbers are used to create simulated data sets in order to study the accuracy of mathematical approximation and the effects of assumptions being violated. 

To generate normal random numbers, we will use `rnorm` function.

There are three arguments in `rnorm`. The first one specifies the number of random numbers to generate, and the second and third are the mean and standard deviation, which, by default, are 0 and 1. For example,
```{r}
rnorm(5)      # generate 5 numbers from standard normal dist.
rnorm(5,5,2)  # generate 5 numbers from a normal dist. with mean=5 and sd=2.
```

##### <span style="color:purple"> Exercise 6. Normal Random Numbers.</span>

(a) Generate 6 random numbers from $N(100,5)$.
(b) Generate 5 random numbers from standard normal distribution.
(c) Generate 10 random numbers from a normal distribution that is centered at -18 with variance equal to 16. 

