---
title: "STAT 1060 Data Science Foundations"
subtitle: "Lab 7 Two-Sample t procedures and Permutation Tests"
author: "Student Version"
date: "11/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, out.width="70%", cache=T)
```

# Introduction

```{r}
file.choose('Anorexia.txt')
```

```{r}
Anorexia <- read.table('/Users/SanjanaPejathaya/Downloads/Anorexia (1).txt', header = TRUE)
dim(Anorexia)
```

```{r}
Anorexia[sample(71,5),]
```

```{r}
unique(Anorexia$Therapy)
```

# Two-Sample T-test

Let $\mu_{F}$ and $\mu_{C}$ be the mean weight gain of the "Family" and "Control"
treatments respectively. The hypothesis test we want to perform is as follows.

\begin{equation}
\begin{aligned}
H_{0}: & \mu_{F} = \mu_{C} \\
H_{a}: & \mu_{F} > \mu_{C}
\end{aligned}
\end{equation}

or similarly stated

\begin{equation}
\begin{aligned}
H_{0}: & \text{ The true mean weight gain is equal between the Family and Control
treatments} \\
H_{a}: & \text{ The true mean weight gain is greater on the Family treatment than 
the Control}
\end{aligned}
\end{equation}

```{r}
Gains.Fam <- Anorexia$Y[ Anorexia$Therapy == "Family" ]
Gains.Cont <- Anorexia$Y[ Anorexia$Therapy == "Control" ]
```

```{r}
t.test(Gains.Fam, Gains.Cont, alternative="greater")
```

# Permutation test

```{r}
perm_fun <- function(x, nA, nB)
{
n <- nA + nB
idx_b <- sample(1:n, nB)
idx_a <- setdiff(1:n, idx_b)
mean_diff <- mean(x[idx_b]) - mean(x[idx_a])
return(mean_diff)
}
```

```{r}
Y1 = Anorexia$Y[ Anorexia$Therapy != "Cog/Behav"]
Gains.Fam <- Anorexia$Y[ Anorexia$Therapy == "Family" ]
Gains.Cont <- Anorexia$Y[ Anorexia$Therapy == "Control" ]
nA = length(Gains.Fam)
nB = length(Gains.Cont)
```

```{r}
set.seed(100)
# Call the perm_fun 1000 times to get 1000 mean differences
perm_diffs <- rep(0, 1000)
for (i in 1:1000) {
perm_diffs[i] = perm_fun(Y1, nA, nB)
}
# Calculate the "p-value"
mean_diff = mean(Gains.Fam)-mean(Gains.Cont)
mean(perm_diffs > mean_diff)
```


