---
title: "STAT 1060 Data Science Foundations"
subtitle: "Homework 7"
#author: "Instructor Version"
#date: "8/12/2022"
output: 
    #html_document
    word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.
* Due date and time: 11:59 PM, Friday, November 17, 2023.

#### Questions 1-4: Exercises in Lab 6 and Lab 7
The primary diagnosis for anorexia is a weight that is less than 85% of what is considered normal for that person's height and age. In this population of patients, researchers consider a patient anorexic if his or her weight is lower than 90 lbs. 

Read this data set into your R session:
```{r}
Anorexia <- read.table("/Users/SanjanaPejathaya/Downloads/Anorexia (1).txt", header=TRUE)
dim(Anorexia)
```

The columns `Before` and `After` are the weights of the patients before and after the treatments, respectively. The last column `Y` is the difference of weights: `After - Before`. 

#### <span style="color:purple"> Exercise 1. (5 points) One-Sample t-Test  </span>
Question: Is there evidence that the patients are anorexic after treatment? Let $\alpha=0.05$.

```{r, echo = T}
Fam <- Anorexia[Anorexia$Therapy == "Family", ]
dim(Fam)

t.test(Fam$Y, alternative="greater", mu = 0, conf.level = 0.95)

```

#### <span style="color:purple"> Exercise 2. (10 points) Paired t-Test  </span>
Question: Was the average weight gain in the `Cog/Behav` group large enough to provide evidence that the `Cog/Behav` treatment worked? Let $\alpha=0.05$.

```{r, echo=T}
t.test(Fam$After, Fam$Before, alternative="greater", paired = TRUE)
```

#### <span style="color:purple"> Exercise 3. (10 points) Two-Sample t-Tests  </span>
Test the following hypotheses:

* H0: "The true mean weight gain is equal between the `Cog/Behav` and `Control` treatments" 
* Ha: "The true mean weight gain is greater on the `Cog/Behav` treatment than the Control". 

```{r, echo=T}
CogBehav.Fam <- Anorexia$Y[ Anorexia$Therapy == "Cog/Behav" ]
Control.Cont <- Anorexia$Y[ Anorexia$Therapy == "Control" ]

t.test(CogBehav.Fam, Control.Cont, alternative="greater")

```
The test is significant since the p-value is 0.05232. We can conclude that the Cog/Behav treatment is effective at significance level, say, 0.01.

#### <span style="color:purple"> Exercise 4. (10 points)  Permutation A/B Test  </span>

To apply a permutation test, we need a function to randomly assign the 36 times to the two groups.

```{r}
perm_fun <- function(x, nA, nB)
{
  n <- nA + nB
  idx_b <- sample(1:n, nB)
  idx_a <- setdiff(1:n, idx_b)
  mean_diff <- mean(x[idx_b]) - mean(x[idx_a])
  return(mean_diff)
}
```

```{r, echo = T}
Y1 = Anorexia$Y[ Anorexia$Therapy != "Cog/Behav"]
Gains.Fam <- Anorexia$Y[ Anorexia$Therapy == "Family" ]
Gains.Cont <- Anorexia$Y[ Anorexia$Therapy == "Control" ]
nA = length(Gains.Fam)
nB = length(Gains.Cont)

```

Calling this function R = 1,000 times leads to a distribution of differences. 

Test the following hypotheses using a permutation test:

* H0: "The true mean weight gain is equal between the `Cog/Behav` and `Control` treatments" 
* Ha: "The true mean weight gain is greater on the `Cog/Behav` treatment than the Control".

Use significance level $\alpha=0.01$. Set the seed to be 99.

```{r, echo = TRUE}

set.seed(100)
# Call the perm_fun 1000 times to get 1000 mean differences
perm_diffs <- rep(0, 1000)
for (i in 1:1000) {
perm_diffs[i] = perm_fun(Y1, nA, nB)
}
# Calculate the "p-value"
mean_diff = mean(Gains.Fam)-mean(Gains.Cont)
mean(perm_diffs > mean_diff)

```


#### <span style="color:purple"> Question 5. (15 points) Two-Sample t-Tests (Pooled)  </span>
(a) (5 points) Create a Q-Q plot for `Y` in the `Family` group and a Q-Q plot for the `Control` group. Are they approximately normal?

```{r, echo=TRUE}
#family_data <- subset(Anorexia, Group == "Family")

qqnorm(Fam$Y)
qqline(Fam$Y, col = 2) 

# Label the plot
title("Q-Q Plot for Family Group")

#control_data <- subset(Anorexia, Group == "Control")

# Create Q-Q plot for the 'Control' group
qqnorm(Control.Cont)
qqline(Control.Cont, col = 2)  # Add a reference line

# Label the plot
title("Q-Q Plot for Control Group")
```

(b) (5 points)  Conduct a equal-variance test using the `var.test` function. Can we conclude that the family group and control group have different variance?

```{r, echo=T}
var_test_result <- var.test(Fam$Y, Control.Cont)

print(var_test_result)

```
(c) (5 points)  Conduct a pooled t test see whether there is significant evidence that the family treatment works. Let $\alpha=0.01$.

```{r, echo =T}

pooled_t_test_result <- t.test(Fam$Y, Control.Cont, var.equal = TRUE)

# Print the result
print(pooled_t_test_result)

```
Since the p-value is less than the chosen significance level, 0.002491<0.01, there is significant evidence that we can reject the null hypotheses, suggesting that the family treatment works. 
