---
title: "STAT 1061 Data Science Foundations"
subtitle: "Homework 5"
#author: "Instructor Version"
#date: "8/12/2022"
output: 
    #html_document
    word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.
* Due date and time: 11:59 PM, Monday, October 30, 2023.

#### Questions 1-3: Exercises 1-3 in Lab 5

a. Generate 1000 observations from `Uniform(0,1)`. This is equivalent to 1000 samples of size 1. This shows the distribution of the population. 
b. Generate 1000 random samples of sizes 10, 30, and 100 respectively. 
c. Calculate the sample mean of each random sample for each sample size.
d. Create a histogram of the sample means for each sample size.

```{r,out.width="80%"}
par(mfrow=c(2,2))
m=1000              # Specify the number of samples
x_bar=matrix(0,4,m) # x_bar will be used to store all the sample means

n=c(1,10,30,100)      # Sample sizes are 10, 30, and 100. 
for(i in 1:4){
  for(j in 1:m){
    x_bar[i,j]=mean(runif(n[i]))
  }
  hist(x_bar[i,],ylab="Relative Frequency",freq=F,main=paste("n=",n[i]),xlab="sample mean")
}
```

#### <span style="color:purple"> Exercise 1. (5 points) </span>
Verify that the mean of sample means is close to the population mean. 

``` {r, echo = TRUE}
mean(x_bar)
```
 


#### <span style="color:purple"> Exercise 2. (5 points)  </span>
Calculate the standard error of sample means and compare it with the theoretical value $\sigma_x/\sqrt{n}$.

Hint: The variance of Unif(0,1) is 1/12.

``` {r, echo = TRUE}
population_variance <- 1/12

se_theoretical <- sqrt(population_variance)

se_sample_means <- sqrt(population_variance / n)

se_theoretical
se_sample_means

```

#### <span style="color:purple"> Exercises 3. (30 points) Exponential and Binomial Distributions </span>
Now, for the Exp(1) and Binom(10,0.1) distributions, 

a. Simulate 1000 random numbers with 4 sample sizes (1, 10, 30, 100)
b. Calculate sample means
c. Create histograms of the sample means
d. Calculate mean and standard errors of sample means.

To generate $n$ random numbers from Exp(1), use `rexp(n)`. 
To generate $n$ random numbers from Binom(10,0.1), use `rbinom(n,10,0.1)`.

``` {r, echo = TRUE}
set.seed(123)
num_samples <- 1000

sample_sizes <- c(1, 10, 30, 100)

exp_sample_means <- matrix(0, nrow = length(sample_sizes), ncol = num_samples)
for (i in 1:length(sample_sizes)) {
  for (j in 1:num_samples) {
    exp_sample_means[i, j] <- mean(rexp(sample_sizes[i], rate = 1))
  }
}

binom_sample_means <- matrix(0, nrow = length(sample_sizes), ncol = num_samples)
for (i in 1:length(sample_sizes)) {
  for (j in 1:num_samples) {
    binom_sample_means[i, j] <- mean(rbinom(sample_sizes[i], size = 10, prob = 0.1))
  }
}


par(mfrow = c(2, 2))

for (i in 1:length(sample_sizes)) {
  # Histogram for Exp(1) distribution
  hist(exp_sample_means[i,], ylab = "Relative Frequency", freq = F, main = paste("Exp(1), n =", sample_sizes[i]), xlab = "Sample Mean")
  
  mean_exp <- mean(exp_sample_means[i,])
  se_exp <- sd(exp_sample_means[i,]) / sqrt(sample_sizes[i])
  
  cat("Exp(1), Sample Size n =", sample_sizes[i], "\n")
  cat("Mean of Sample Means:", mean_exp, "\n")
  cat("Standard Error:", se_exp, "\n\n")
}

for (i in 1:length(sample_sizes)) {
  # Histogram for Binom(10, 0.1) distribution
  hist(binom_sample_means[i,], ylab = "Relative Frequency", freq = F, main = paste("Binom(10, 0.1), n =", sample_sizes[i]), xlab = "Sample Mean")
  
  # Calculate mean and standard error for Binom(10, 0.1) distribution
  mean_binom <- mean(binom_sample_means[i,])
  se_binom <- sd(binom_sample_means[i,]) / sqrt(sample_sizes[i])
  
  cat("Binom(10, 0.1), Sample Size n =", sample_sizes[i], "\n")
  cat("Mean of Sample Means:", mean_binom, "\n")
  cat("Standard Error:", se_binom, "\n\n")
}

```


#### <span style="color:purple"> Exercises 4. (10 points) </span>
Visualize the Law of Large Numbers

(a) Imitate the code in Lab 5 and visualize the LLN using 5000 Exp(1) random numbers. Note that the mean of Exp(1) distribution is 1.

``` {r, echo = TRUE}
set.seed(123)
n <- 5000; x <- rnorm(n); x_bar <- rep(0,n)
for(i in 1:n){x_bar[i] = mean(x[1:i])}
plot(x_bar, type="l"); abline(h=0,col=2,lty=2)

```


(b) Imitate the code in Lab 5 and visualize the LLN using 5000 Unif(-5,5) random numbers. Note that the mean of Unif(-5,5) is 0.

``` {r, echo =TRUE}
set.seed(123)
num_samples <- 5000
random_numbers <- runif(num_samples, min = -5, max = 5)
sample_means <- cumsum(random_numbers) / (1:num_samples)
plot(1:num_samples, sample_means, type = "l", col = "blue", 
     xlab = "Sample Size", ylab = "Sample Mean", 
     main = "Law of Large Numbers (LLN) Visualization")
abline(h = 0, col = "red", lty = 2)


```
