---
title: "STAT 1060 Data Science Foundations"
subtitle: "Homework 2 (Fall 2023)"
#author: "Instructor Version"
#date: "8/12/2022"
output: 
    #html_document
    word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.

Use the following command to load the data into your R session:
```{r}
babies <- read.table("http://www.math.montana.edu/shancock/data/Bwt.dat", header=TRUE, sep=",")
```

The variables are

- `bwt`: baby's weight in ounces at birth
- `gestation`: duration of pregnancy in days
- `parity`: parity indicator (first born = 1, later birth = 0)
- `age`: mother's age in years
- `height`: mother's height in inches
- `weight`: mother's weight in pounds (during pregnancy)
- `smoke`: indicator for whether mother smokes (1=yes, 0=no)

Check that the data were read in correctly:
```{r}
head(babies,3)
dim(babies)	
is.data.frame(babies)

```
#### Questions 1-6: Collect your solutions to Exercises 3-8 in Lab 2. 

##### Exercise 3. (5 points) Order the `babies` data set by mother's smoking status and age.
```{r, echo=TRUE}
order_age = order(babies$age)
babies1 = babies[order_age,]
head(babies1)

order_smoke = order(babies$smoke)
babies1 = babies[order_smoke,]
head(babies1)

```
##### Exercise 4. (5 points) Sort the `babies` data set by mother's height in descending order.

``` {r, echo = TRUE}
sort_height = order(-babies$height)
babies1 = babies[sort_height,]
head(babies1)


```


##### Exercise 5. (5 points) Create a new data set which only contains two columns: `height` and `weight`. Call it `babies_sub2`.

``` {r, echo = TRUE}

babies_sub2 = babies[,c(5,6)]
head(babies_sub2)
```


##### Exercise 6. (5 points) Select all the observations with mother's height that is greater than 68 inches.
``` {r, echo = TRUE}
babies_height = babies[babies$height > 68,]
head(babies_height)

```

##### Exercise 7. (5 points) Calculate the five-number summary of `bwt` in the babies dataset.

``` {r, echo = TRUE}

quantile(babies[['bwt']])

```

##### Exercise 8. (5 points) Find the mean and the median of `weight`. Are they close to each other?

```{r, echo = TRUE}
mean(babies[['weight']])
median(babies[['weight']])
```
Yes. The mean and median are quite close together, with a difference of 342787 units. 


#### Question 7 (4 points) Create a horizontal boxplot for `bwt`

``` {r, echo = TRUE}

boxplot(babies[['bwt']], horizontal = TRUE, data = babies[['bwt']], main = "Baby's Weight In Ounces At Birth, (bwt)", xLab = "bwt", xlab = 'bwt', col = 'pink')

```

#### Question 8 (6 points) Calculate 10% trimmed mean and the median absolute deviation from the median for `bwt`.

``` {r , echo = TRUE}
mean(babies[['bwt']], trim=0.1)
mad(babies[['bwt']])
```

#### Question 9 (5 points) Create a frequency table for `bwt`.
Requirements:

* There should be 15 intervals
* The intervals should be left-closed

``` {r, echo = TRUE}



breaks = seq(from=min(babies1[['bwt']]/1e6),
to=max(babies1[['bwt']])/1e6, length=15)
bwt_binned = cut(babies1[['bwt']]/1e6, breaks=breaks,
left=TRUE, # indicating the intervals should be closed on the right (and open on the left)
include.lowest=TRUE)
as.data.frame(table(bwt_binned))

```


#### Question 10 (5 points) Create a histogram for `bwt`.

Requirements:

1. Use the `breaks` defined in Question 9
2. Give a title "Baby Weights".
3. The x-axis label should be "Baby Weight (in ounce)"

```{r, echo=TRUE}
hist(babies1[['bwt']]/1e6, breaks=breaks,
xlab='Baby Weight (in ounce)',
main='Baby Weights', col = 'pink')


```



