---
title: "STAT 1060 Data Science Foundations"
subtitle: "Homework 4 (Solution)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.

#### Questions 1-6: Exercises 1-6 in Lab 4.

| Type | R Function | Example|
|:--|:--|:--|
|density function | `dnorm(x=, mean=0, sd=1)` | `dnorm(1.5)`|
|cumulative distribution function | `pnorm(q=, mean=0, sd=1, lower.tail = TRUE)`| `pnorm(1.5, 100, 15)`|
|quantile function| `qnorm(p=, mean=0, sd=1, lower.tail = TRUE)` | `qnorm(0.95)`|
|pseudo-random number| `rnorm(n=, mean=0, sd=1)`| `rnorm(100,2,10)`|


##### <span style="color:purple"> Exercise 1. (3 points) Calculate $P(Z>4)$ where $Z\sim N(0,1)$.</span>

``` {r, echo = TRUE}
pnorm(135, mean = 100, sd = 15, lower.tail=FALSE)

```


##### <span style="color:purple"> Exercise 2. (3 points) Calculate $P(85 <X < 115)$ where $X\sim N(100,15).$  </span>

```{r, echo=TRUE}

prob_less_than_85 <- pnorm(85, 100, 15)
prob_less_than_115 <- pnorm(115, 100, 15)
prob_between_85_and_115 <- prob_less_than_115 - prob_less_than_85
print(prob_between_85_and_115)


```

##### <span style="color:purple"> Exercise 3. (6 points) </span>

Typical nightly hours slept by college students is a normal random variable with $\mu = 7$ and $\sigma = 1.5$. 

(a) What is the probability that a student slept less than 6 hours?

```{r, echo=TRUE}
 pnorm(6, 7, 1.5)
```

(b) What is the probability that a student slept more than 12 hours?
```{r, echo=TRUE}
 1 - pnorm(12, 7, 1.5)

```
##### <span style="color:purple"> Exercise 4. (6 points) Standard Normal Quantile.</span>
(a) Suppose $P(Z>z)=0.95$, what is the value of $z$?
```{r, echo = TRUE}
qnorm(0.95, mean = 100, sd = 15)
```

(b) What is the $99^{th}$-percentile (or 0.99-quantile) of standard normal distribution?
```{r, echo = TRUE}
qnorm(0.99, mean = 100, sd = 15)


```

##### <span style="color:purple"> Exercise 5. (3 points) Normal Quantiles.</span>
In order to be among the top 1% smartest people, what IQ score one needs to have? 
```{r, echo = TRUE}
qnorm(0.99, 100, 15)

```
##### <span style="color:purple"> Exercise 6.(9 points) Normal Random Numbers.</span>

(a) Generate 6 random numbers from $N(100,5)$.
(b) Generate 5 random numbers from standard normal distribution.
(c) Generate 10 random numbers from a normal distribution that is centered at -18 with variance equal to 16. 

```{r, echo= TRUE}
rnorm(6, 100, 5)
rnorm(5)
rnorm(10,-18,sqrt(16))

```


##### <span style="color:purple"> Question 7. (10 points) Uniform Distributions</span>
Uniform distribution family also has a set of R functions:

| Type | R Function | Example|
|:--|:--|:--|
|density function | `dunif(x, min = 0, max = 1)` | `dunif(0.5)`|
|cumulative distribution function | `punif(q, min = 0, max = 1, lower.tail=TRUE)`| `punif(1.5, 1, 2)`|
|quantile function| `qunif(p, min = 0, max = 1, lower.tail=TRUE)` | `qunif(0.95, 0, 2)`|
|pseudo-random number| `runif(n, min = 0, max = 1)`| `runif(100)`|

Suppose a uniform distribution has the range $(0,2)$. Suppose a random variable $Y\sim Unif(0,2).$

(a) (1 point) By the definition of a density curve, what is the height of the density curve?
The height of a uniform distribution that ranges between a and b, is 1b−a.
(b) (3 points) Find $P(Y\ge 1.2)$.
(c) (3 points) Find $P(0.1<Y<1)$.
(d) (3 points) Find $P(Y\le 0.15)$.

``` {r, echo = TRUE}
punif(1.2, 0, 2, lower.tail=FALSE)
dunif(0.1, 0, 2, F) - dunif(1,0,2,T)
1 - punif(0.15, min = 0, max = 2, lower.tail = FALSE)
```
##### <span style="color:purple"> Question 8. (10 points) Mean and Variance of Random Variables</span>
Suppose that $X$ is a random variable with mean 30 and standard deviation 4. Also suppose that $Y$ is a random variable with mean 50 and standard deviation 8. Assume that the correlation between $X$ and $Y$ is 0.5. 

(a) (2 points) Find the mean of random variable $Z$ where $Z=5X-3Y$.
(b) (2 points) Find the mean of random variable $W$ where $W=10X-35$.
(c) (4 points) Find the variance of random variable $Z$ where $Z=5X-3Y$.
(d) (2 points) Find the variance of random variable $W$ where $W=10X-35$.

```{r, echo=TRUE}
# μx+y = μx + μy
5 *30 + 3 * 50 + 
10 * 30 - 35 




```








