---
title: "Homework Assignment 1: R Basics" 
subtitle: "STAT 1060: Data Science Foundations (Fall 2023)"
output:
#  pdf_document: default
#    fig_height: 5
#    fig_width: 7
#   html_document:
#    fig_height: 5
#    fig_width: 7
   word_document:
#    fig_height: 3
#     fig_width: 5
---

```{r, setup, include=FALSE}
require(mdsr)   # Load additional packages here 
require(tidyr)
require(tm)
require(wordcloud)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


### Instruction

* Total points: 50
* Make sure you show your code as well as the output of your code.
* Due Date and Time: 11:59 PM, Friday, September 15, 2023.

### 1. Assignments (10 points) 
##### 1) (2 points) Assign the letter "A" to a variable called `x`.
##### 2) (2 points) Assign the value 2022 to a variable called `year`.
##### 3) (2 points) Assign the letter logic value "TRUE" to a variable called `y`.
##### 4) (2 points) Let a new variable `z` be equal to `x`. Show the value of z.
##### 5) (2 points) Try to add `year` and `y` together. What do you get?


``` {r,echo =TRUE}
x = "A"; x
year = 2022; year
y = TRUE; y
z = x; z
year + y 
```

### 2. Vectors (10 points)
##### 1) (3 points) Create a vector as follows:
```{r,echo=FALSE}
c("Statistics", "Math", "CS", "Data Science")

```


```{r, echo = TRUE}
subjects = c('Statistics', 'Math', 'CS', 'Data Science'); subjects
```



##### 2) (3 points) Create a vector as follows using `seq()`:
```{r,echo=FALSE}
seq(-100,0,10)
```
```{r, echo=TRUE}
seq(-100,0, 10)
```
##### 3) (4 points) Create a vector as follows using `rep()`:
```{r,echo=FALSE}
rep(c('A','C','D'), 3)
```
```{r,echo=TRUE}
v = c("A","C","D")
rep(v,3) 
```

### 3. Errors (10 points)
For each of the following assignment statements, describe the error (or note why it does not generate an error).

```{r, eval=FALSE}
result1 = log 10
result2 == "Hello to you!"
3result <- "Hello to you"
result4 <- Hello to you
result5 <- date()
```
result1 = log 10
  It states an unexpected numeric constant 
result2 == "Hello to you!"
  Error: object 'result2' not found
3result <- "Hello to you"
  Error: unexpected symbol in "3result"
result4 <- Hello to you 
  Error: unexpected symbol in "result4 <- Hello to"
result5 <- date()
  This line worked because the variable was defined correctly. 

### 4. Matrices (10 points)
##### 1) (4 points) Create matrices as follows by writing code. Call them `M1` and `M2`, respectively. 
\ \  
```{r,echo=FALSE}
a <- 1:12
M1 <- matrix(a,3,4)
print(M1)
M2 <- matrix(a,3,4,byrow=TRUE)
print(M2)
```

```{r, echo= TRUE}
M1 = matrix(1:12,3,4); M1

M2 = matrix(1:12,3,4, byrow = T); M2
```

##### 2) (2 points) Extract the value at the 2nd row and 4th column of `M1`.

    # value is 11

##### 3) (2 points) Extract the values in the 1st row in `M1`. 

    # values from 1st row, M1 = (1,4,7,10)

##### 4) (2 points) Extract the values in the 2nd column in `M2`.

    # values from the 2nd column, M2 = (2,6,10)

### 5. Data Frames (10 points)
##### 1) (2 points) Create a data frame as follows:
```{r,echo=FALSE}
Name <- c("Alice","Betty","Caitlyn","Donna","Ella")
Year <- c("1","2","2","4","3")
GPA <- c(3.41, 3.53,3.32,3.67,4.00)
Students <- data.frame(Name,Year,GPA)
Students
```

```{r, echo=TRUE}
Name <- c("Alice", "Betty", "Caitlyn", "Donna", "Ella")
Year <- c(1, 2, 3, 4, 3)
GPA <-c(3.41, 3.53, 3.32, 3.67, 4.00)

df <- data.frame(Name, Year,GPA)

print (df)

```

##### 2) (2 points) Use column name to extract the names of students. 

Column name has been defined as "Name"

##### 3) (2 points) Select the information about Ella.

 Ella is in her 3rd year with a GPA of 4.00.

##### 4) (2 points) List name(s) of students whose GPA is lower than 3.5.

* Alice
* Caitlyn

##### 5) (2 points) List the GPA of all 3rd-year students.

* Caitlyn 
  GPA = 3.32 GPA





